---
title: "Pulling microcontroller data from google spreadsheets"
output:
  html_document:
    code_folding: show
    df_print: paged
    number_sections: no
    toc: yes
    toc_float: yes
editor_options:
  chunk_output_type: inline
---

# Install prereqs

```{r, eval=FALSE}
# R
install.packages(c("googlesheets", "lubridate"))
# python
source activate class
pip install gspread
```

# Pull data in R

```{r setup, include=FALSE}
library(tidyverse)
library(googlesheets)
library(lubridate)
library(plotly)
library(reticulate)
# load python
use_condaenv("class", conda = "auto", required = TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

## Authenticate

When you run `gs_auth()` for the first time (easiest from within RStudio rather than in the middle of a knitted doc!), it asks you for your google credentials and stores them in an `.httr-oauth` file. Delete this file to remove your credentials and NEVER commit an .httr-oauth file to github - best to always include the following lines in your `.gitignore` file:

```
# exclude authentication tokens
.httr-oauth
```

```{r}
gs_auth()
```

## Identify spreadsheet

```{r}
# identify spreadsheet by name
spreadsheet_title <- "class log test"
gs <- gs_title(spreadsheet_title)
```

## Read data from worksheet (tab)

```{r}
data <- gs %>% gs_read_csv(ws = "light")
```

## Process data

The key is to adjust the timezone from the UTC based timestamp. Notice a couple of data outages.

```{r}
data %>% 
  mutate(published_at = published_at %>% with_tz("America/Denver")) %>% 
  ggplot() +
  aes(x = published_at, y = cds, color = location) +
  geom_line() +
  scale_x_datetime() +
  theme_bw()
```

Overplot individual days. Notice the shift in daylight savings that suddenly moved everything by an hour!

```{r}
data %>% 
  mutate(
    published_at = published_at %>% with_tz("America/Denver"),
    day = paste(month(published_at, label = TRUE), day(published_at)),
    time = published_at %>% { hour(.) + minute(.)/60 + second(.)/3600 }
  ) %>% 
  ggplot() +
  aes(x = time, y = cds, color = day) +
  geom_line() +
  scale_x_continuous(breaks = 0:24, expand = c(0,.01)) +
  theme_bw()
```

Interactive plot.

```{r}
ggplotly()
```


# Pull data in Python

Use the [gspread](https://gspread.readthedocs.io/en/latest/) library. 

```{python}
import gspread
```

