---
title: "Numerical computation"
output:
  html_document:
    code_folding: show
    df_print: paged
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
library(tidyverse) # key packages are dplyr and tidyr
library(reticulate)
# load python
use_condaenv("class", conda = "auto", required = TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

# Basics

## Python

The `numpy` library provides the basis for numerical multidimensional arrays. [`xarray`](http://xarray.pydata.org/en/stable/) provides additional features that allows for numpy arrays with pandas-like named dimensions and therefore very efficient representation of netCDF data and similar geospatial information.

## R

All basic data in R is always already a vector (i.e. a 1-dimensional array). Two dimensional number arrays in R are called `matrix` but the `array` data type provides more general multi-dimensional data options similar to `numpy` in python. However, the more common approach in R is to use `data_frame` tabular data that can be nested (tidyr's `nest`) and unnested (tidyr's `unnest`) to provide higher dimmensional data that's still compatible with `dplyr` and `ggplot` operations. 

# Create arrays

## Python

```{python}
# python
import numpy as np # import numpy

# basic numpy array (by default initialized as float)
a = np.array([3.14, 60, 17])
print(a)
print(type(a[0]))

# specify integer arrays if desired
b = np.array([1, 2, 3], dtype=int)
print(b)
print(type(b[0]))
```

```{r, message=FALSE}
# basic vector
a <- c(3.14, 60, 17) 
str(a)
# integer vector
b <- c(1L, 2L, 3L) 
str(b)
# special case for easy creation of sequential integer vectors
b2 <- 1:5
str(b2)
```


```{python}
# python
# generate a basic array of zeros
z = np.zeros((2, 5, 10)) # 2 x 5 x 10 array
print(z)

# generate an array of ones with same dimension
z2 = np.ones_like(z)
print(z2)

# range of 100 values from 0.1 to 10 reshaped into a 10x10 array
z3 = np.linspace(0.1,10,100).reshape(2, 5, 10)
print(z3)
```

## R

```{r}
# R
# notice that although the dimensions are the same size, the way the arrays are populated is quite different!
str(py$z3) # array coming from python
z3 <- array(seq(0.1, 100, by = 0.1), dim = c(2, 5, 10))
str(z3) # "same" array created in R

# default array values - either e.g. a sequence as illustrated above or single default
str(array(NA_real_, dim = c(2, 5, 10))) # not available as default
str(z <- array(0L, dim = c(2, 5, 10))) # 0 integer as default
str(z2 <- array(1, dim = c(2, 5, 10))) # 1.0 numeric (=float) as default

# note that arrays in R can also have named dimensions
zdim <- array(1:100, dim = c(2, 5, 10), dimnames = list(x = str_c("x", 1:2), y = str_c("y", 1:5), z = str_c("z", 1:10)))
str(zdim)
print(zdim)
```

# Structure information

## Python

```{python}
# python
# get shape and size information
print(np.shape(a)) # dimensions
print(np.size(a)) # total # elements
print(np.shape(z)) # dimensions
print(np.size(z)) # total # elements
# reshaping into different dimensions
print(z3.reshape(20,5))
```

## R

```{r}
# R
# get shape and size information
print(dim(a)) # dimension is NOT defined for a vector
print(length(a)) # total # elements
print(dim(z)) # dimensions of array
print(length(z)) # total # elements
# reshaping into different dimensions
z3_copy <- z3
str(z3_copy)
dim(z3_copy) <- c(20, 5)
str(z3_copy)
```


# Selectors

## Python

```{python}
# python
# select subsets of the array by 0-based index
print(a)
print(a[0]) # index from beginning
print(a[0:2]) # range (EXCLUSIVE end index)
print(a[-2]) # index from end
print(a[-2:]) # range till end
# same for multidimensional arrays, each dimension as separate index
print(z3[0:2,3:,1:4])
```

## R

```{r}
# R
# quite different from python! indexing is 1-based and the end index is INCLUSIVE, also the -x syntax means exclude, not count from the end
print(a)
print(a[1]) # index from beginning (1-base)
print(a[1:2]) # range (INCLUSIVE end index)
print(tail(a, 2)) # get as indexed from back
print(a[-2]) # EXCLUDE this index
print(a[-c(2:3)]) # EXCLUDE range
# same for multidimensional arrays, each dimension as separate index but 1-based
str(z3[1:2,3:4,2:4])
```

# Filtering

## Python

```{python}
# python
# use where for filtering
[i] = np.where(a < 20) # get indices (0-based as always)
print(i)
print(a[i])
print(a[np.where(a < 20)]) # directly
# same for multidimensional arrays except multiple indices are returned
[i,j,k] = np.where(z3 > 9.5) # store indices
print(z3[i,j,k])
print(z3[np.where(z3 > 9.5)]) # directly
```

## R

```{r}
# R
i <- which(a < 20) # get indices (1-based as always)
print(i)
print(a[i])
print(a[a < 20]) # directly
# same for multidimensional arrays
is <- which(z3 > 9.5) # different to python! single index vector returned (array is still just a vector underneath)
print(z3[is]) # consequently selected values no longer have their dimensionality!
print(z3[z3 > 9.5]) # directly
```

## R alternative: switch to data frames

```{r}
# R: alternative
# this is where conversion to data frame becomes more and more useful
array_to_df <- function(array) { 
  array %>% 
    # convert array to table (+ to dplyr data frame)
    as.data.frame.table() %>% tbl_df() %>%
    # switch to index instead of named dimensions
    mutate_if(is.factor, as.integer) %>% 
    return()
}
  
# convert array to data frame (note: column renaming is optional)
z3_df <- z3 %>% array_to_df() %>% rename(x = Var1, y = Var2, z = Var3, value = Freq)
z2_df <- z2 %>% array_to_df() %>% rename(x = Var1, y = Var2, z = Var3, value = Freq)

# what does this look like?
z3_df

# now both selection and filtration all become the same process & very intuitive
z3_df %>% filter(x %in% 1:2, y %in% 3:4, z %in% 2:4) # by 'index'
z3_df %>% filter(value > 9.5) # by value, all index info also preserved

# nesting / unnesting: easy way to step up/down the dimensionality
z3_df %>% nest(-x) # nest everything but the x dimension
z3_df %>% nest(-x, -y) # nest everything but x and y
z3_df %>% nest(-x) %>% unnest(data) # back where we started
```


# Calculations

## Python

```{python}
# python
# are performed element-wise!
z_plus = z2 + z3
print(z_plus)
z_times = z2 * z3
print(z_times)
```

```{python}
# better to be explicit using the .multiply / .dot functions
print(np.multiply(z2, z3)) # element-wise multiplication

# dot product does not work with these dimensions
try:
  print(np.dot(z3, z3)) 
except:
  print("Error:", sys.exc_info()[1])
  
# try again with compatible dimensions
print(np.dot(z3.reshape(10, 10), z3.reshape(10, 10))) 
```

## R

```{r}
# R
# are also performed element-wise by default
z_plus = z2 + z3
str(z_plus)
z_times = z2 * z3
str(z_times)

# in data frame format
z2_df %>% 
  # combine with z3 data frame that has the same dimensions
  full_join(z3_df, by = c("x", "y", "z"), suffix = c(".z2", ".z3")) %>% 
  # make any element by element calculations
  mutate(
    z_plus = value.z2 + value.z3,
    z_times = value.z2 * value.z3
  )

# matrix multiplication is a bit trickier, a few examples
(v <- 1:4)
v %*% v
(w <- diag(v))
w %*% v
(m <- array(1:12, dim = c(4,3)))
v %*% m
tryCatch(m %*% m, error = function(e) message("Error: ", e$message))
```

# Logical Operations

## Python

```{python}
# python - element wise logic
print(a < 20)
# logic on multiple values
print(np.logical_or(a < 4.0, a > 20.0))
# find out what the values are of the elements that match this condition
print(a[np.logical_or(a < 4.0, a > 20.0)])
# logic for multidimensional arrays
print(z3[np.logical_or(z3 < 0.5, z3 > 9)])
```

## R

```{r}
# R - also element wise logic
print(a < 20)
print(a < 4.0 | a > 20.0) # since everything is a vector, this works directly with vector logic (&, |)
print(a[a < 4.0 | a > 20.0])
# logic for multidimensional arrays
print(z3[z3 < 0.5 | z3 > 9])

# or the data frame version
z3_df %>% filter(value < 0.5 | value > 9)
```


# Numerical calculation

## Example 1: radioactive decay with production term

$$
\frac{dC}{dt} = -kC + P
$$

Analytical solution

$$
C(t) = \frac{P}{k}\left(1-e^{-kt}\right)
$$

Numerical solution using finite difference.

$$
\frac{dC}{dt} \approx \frac{C^{t+1} - C^t}{\Delta t} = -kC^t + P \\
C^{t+1} = C^t - k \Delta t C^t + P \Delta t
$$

## Python

```{python}
# python - define parameters
C = 0.0
dt = 100.0
run_duration = 2.0e6
num_time_steps = int(run_duration / dt)
hl = 730000
k = np.log(2.0)/hl
P = 20.0
C = np.zeros(num_time_steps)
```


```{python}
# python - run model
for i in range(1,num_time_steps):
  C[i] = C[i-1] + dt * (-k * C[i-1] + P)
```

```{python}
# python - plot model result
import matplotlib.pyplot as plt
t = np.arange(0.0, run_duration / 1000.0, dt / 1000.0)
plt.plot(t, C)
plt.xlabel("Time [ky]")
plt.ylabel("C")
plt.show()
```

## R

```{r}
# R - define parameters 
# could do this in global variables but more common would use a data frame
params <- data_frame(
  C0 = 0.0,
  dt = 100.0,
  run_duration = 2.0e6,
  num_time_steps = as.integer(run_duration / dt),
  hl = 730000,
  k = log(2.0)/hl,
  P = 20.0
)
params
```

```{r}
# R - initialize data outline
data <- data_frame(
  # time axis (in 1000 years)
  time.ka = seq(0, by = params$dt/1000, length.out = params$num_time_steps),
  # data with lots of NAs
  C = c(params$C0, rep(NA, params$num_time_steps - 1))
)
data

# run the numerical model
for (i in 2:params$num_time_steps) {
  data$C[i] <- data$C[i-1] + params$dt * (-params$k * data$C[i-1] + params$P)
}
```

```{r}
# R - plot
data %>% 
  ggplot() + 
  aes(x = time.ka, y = C) + 
  geom_line() +
  theme_bw() +
  labs(x = "Time (ky)", y = "C") 
```


