---
title: "Introduction to function documentation, data import, data structures and plotting"
output:
  html_document:
    code_folding: show
    df_print: paged
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options:
  chunk_output_type: inline
---

<!-- {% raw %} -->

```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
library(latex2exp)
library(reticulate)
# load python
use_condaenv("class", conda = "auto", required = TRUE)
knitr::opts_chunk$set(echo = TRUE)
```

# Function Documentation

```{python}
# python
def pythagorize(x, y):
  """
  Calculate and return hypothenuse given x and y
  Parameters
  =========
  x: float
     Distance in x direction
  y: float
     Distance in y direction
  """
  return(x*x + y*y) ** 0.5

print(pythagorize(3, 4))
```


```{r}
# R
#' Calculate and return the hypothenuse given x and y
#' @param x distance in x direction
#' @param y distance in y direction
#' @return the diagonal
pythagorize <- function(x, y) {
  return(sqrt(x^2 + y^2))
}
pythagorize(3, 4)
```


# Import

## Slope Analysis Data

Import the slope analysis data.

```{r}
# R
data_slope_analysis <-
  # file path to slope analysis csv data
  file.path("data", "slope_analysis20180826.csv") %>%
  # read csv data
  read_csv(
    # define column types as safety check
    col_types = cols(
      `Run name` = col_character(),
      `Disturbance rate parameter` = col_double(),
      `Weathering rate parameter` = col_double(),
      `Slope angle` = col_double(),
      `Slope gradient` = col_double()
    )
  ) %>%
  # select and rename relevant columns
  select(
    angle = `Slope angle`,
    wprime = `Weathering rate parameter`,
    dprime = `Disturbance rate parameter`
  ) %>%
  # scale the dimensionless numbers
  mutate(
    wprime = 1e3 * wprime,
    dprime = 1e3 * dprime
  )

# print out data frame
data_slope_analysis
```

## Dissolution Only Data

Import the dissolution only data.

```{r}
# read dissolution analysis csv data
data_dissolution_only <-
  # file path
  file.path("data", "dissolution_only_results20180717.csv") %>%
  # read csv
  read_csv(
    # skip the first row which has landlab info
    skip = 1,
    # define columns for safety
    col_types = cols(
       `Dissolution rate parameter (1/yr)` = col_double(),
       `Gradient (m/m)` = col_double(),
       `Slope angle (deg)` = col_double()
    )
  ) %>%
  # select and rename relevant columns
  select(
    angle = `Slope angle (deg)`,
    dparam = `Dissolution rate parameter (1/yr)`
  ) %>%
  # scale identically to the model output
  mutate(
    wprime = 1e3 * dparam
  )

# print data
data_dissolution_only
```

# Processing

## Analytical solution

Calculate the analytical solution for the dissolution-only case.

```{r}
analytical_dissolution_only <-
  data_frame(
    wprime = seq(
      from = min(data_dissolution_only$wprime),
      to = max(data_dissolution_only$wprime),
      length.out = 100),
    angle = 60 - 360 * wprime / pi
  )
analytical_dissolution_only
```

## Combine data

Combine all the data into a uniform format for easy calculations and visualization. Introduce new information columns `scenario` and `analytical` to make it easy to use grouping and aesthetics later on.

```{r}
all_data <-
  # combine data frames
  bind_rows(
    # each dprime is one scenario
    data_slope_analysis %>% mutate(scenario = as.character(dprime), analytical = FALSE),
    # dissolution only
    data_dissolution_only %>% mutate(scenario = "inf", analytical = FALSE),
    # analytical solution
    analytical_dissolution_only %>% mutate(scenario = "analytical", analytical = TRUE)
  ) %>%
  # focus on columns of interest
  select(analytical, scenario, angle, wprime) %>%
  # sort
  arrange(analytical, scenario)

# print out
all_data
```

# Information

Spit out summary information about the data.

```{r}
# summarize how many data points there are scenario and what the min and max angles are
all_data %>%
  group_by(scenario) %>%
  summarize(
    n_data_points = n(),
    angle_min = min(angle),
    angle_max = max(angle)
  )

# summarize at what value of the dimensionless weathering rate the facet angle first falls below 30 degrees (the effective angle of repose)
all_data %>%
  filter(angle < 30) %>%
  group_by(scenario) %>%
  summarize(lowest_w_below_30deg = min(wprime))
```

# Visualization

Plot the facet angle for each value of $d'$ as a function of weathering parameter ($w'$) and add the dissolution only scenario ($d'\rightarrow \infty$) and analytical solution for the latter. Also include the model's effective 30$^\circ$ angle of repose for reference.

```{r}
# plot the combined data
all_data %>%
  # focus on data in the relevant range
  filter(between(wprime, 0.1, 100)) %>%
  # start plot object
  ggplot() +
  # define top level aesthetics
  aes(x = wprime, y = angle) +
  # add data points (i.e. everything that is not the analytical solution)
  geom_point(
    data = function(df) filter(df, !analytical), # data filter
    mapping = aes(shape = scenario), # aesthetics
    size = 3 # symbol size
  ) +
  # add analytical solution
  geom_line(
    data = function(df) filter(df, analytical), # data filter
    mapping = aes(color = scenario), # aesthetics
    size = 1 # line thickness
  ) +
  # add horizontal line for 30 degrees
  geom_hline(
    data = data_frame(y = 30, scenario = "eff_angle"),
    mapping = aes(yintercept = y, color = scenario), size = 1, linetype = 2
  ) +
  # switch x-axis to log scale
  scale_x_log10(expand = c(0, 0))
```

```{r}
# make the basic plot above interactive
# conversion from ggplot to plotly has some issues but works okay for simple plots
# build plotly (also aesthetics based) from scratch for more complicated plots
ggplotly(p = ggplot2::last_plot(), dynamicTicks = TRUE)
```

## Fancify Plot

```{r, message=FALSE, fig.width=8, fig.height=6}
# customize the aesthetics of the basic plot above
ggplot2::last_plot() +
  # shape scale & latex legend for symbols
  scale_shape_discrete(
    # replace plain text labels with rendered latex
    labels = function(x) {
      case_when(
        x == "inf" ~ "$d' \\rightarrow \\infty$",
        x != "inf" ~ sprintf("$d' = 10^{%.0f}$", log10(as.numeric(x)))
      ) %>% TeX()
    }
  ) +
  # color scale & latex legend for lines
  scale_color_manual(
    # selecting manual colors
    values = c("red", "dark green"),
    # replace plain text labels with rendered latex
    labels = function(x) {
      case_when(
        x == "analytical" ~ "$\\theta = 60° - 360° \\frac{w'}{\\pi}$",
        x == "eff_angle" ~ "$\\theta = 30°$"
      ) %>% TeX()
    }
  ) +
  # x axis scale
  scale_x_log10(
    labels = function(x) sprintf("$10^{%.0f}$", log10(x)) %>% TeX(),
    expand = c(0, 0)) +
  # y axis scale
  scale_y_continuous(labels = function(x) str_c(x, "°")) +
  # add labels
  labs(
    x = TeX("Dimensionless weathering rate parameter, $w'$"),
    y = "Facet angle",
    shape = "symbols",
    color = "lines"
  ) +
  # switch to better default theme
  theme_bw() +
  # add additional customization
  theme(
    text = element_text(size = 18), # font size
    legend.text.align = 0, # legend text alignment
    panel.grid.minor = element_blank() # remove minor grid lines
  )
```

## Save Plot

```{r}
ggsave(filename = "facet_angle_vs_wprime.pdf", plot = ggplot2::last_plot(), width = 8, height = 6)
```


## Alternative visualization

This next figure isn't meant for the paper, but is just here to check out what it looks like if you plot *difference* between fault dip and facet angle instead of the facet angle itself.

```{r, fig.width=8, fig.height=6}
# simply switch out the y aesthetic on the previous plot
ggplot2::last_plot() %+% aes(y = 60 - angle) + labs(y = "60° - facet angle")
```

# Interactive Plot

For more complex interactive plots, it's useful to call plot_ly directly.

```{r, message=FALSE}
# start plotly object
plot_ly(
  x = ~ wprime,
  y = ~ angle
) %>%
  # add points
  add_markers(
    data = filter(all_data, between(wprime, 0.1, 100), !analytical),
    color = ~ str_c("d' = ", scenario),
    marker = list(size = 10)
  ) %>%
  # add lines
  add_lines(
    data = filter(all_data, between(wprime, 0.1, 100), analytical),
    color = ~ scenario,
    line = list(color = "black")
  ) %>%
  # add layout specifications
  layout(
    xaxis = list(title = "Dimensionless weathering rate parameter", type = "log"),
    yaxis = list(title = "facet angle")
  )
```

# Animated Plot

```{r}
# animation frames
frames <-
  data_frame(
    wprime_cutoff = 10^seq(from = -1, to = 2, by = 0.05),
    `w'` = round(wprime_cutoff, 3)
  )

# animated data
animated_data <- all_data %>%
  # focus on data between 0.1 and 100 wprime
  filter(between(wprime, 0.08, 100)) %>%
  # cross with the frames data frame
  crossing(frames) %>%
  # ensure each frame only goes to the cutoff
  filter(wprime <= wprime_cutoff)
```

```{r}
# generate animated plot
# Note: also works by defining the frame aesthetic in ggplot and using ggplotly
  plot_ly(
    x = ~ wprime,
    y = ~ angle,
    frame = ~ `w'`
  )  %>%
  # add points
  add_markers(
    data = filter(animated_data, !analytical),
    color = ~ str_c("d' = ", scenario),
    marker = list(size = 10)
  ) %>%
  # add lines
  add_lines(
    data = filter(animated_data, analytical),
    color = ~ scenario,
    line = list(color = "black")
  ) %>%
  # add layout specifications
  layout(
    title = "<b>Wonderfuly Unnecessarily Animated Plot</b>",
    xaxis = list(title = "Dimensionless weathering rate parameter", type = "log"),
    yaxis = list(title = "facet angle")
  ) %>%
  # animation options
  animation_opts(
    frame = 100,
    transition = 0,
    redraw = FALSE
  )
```

<!-- {% endraw %} -->
