---
title: "Parallel computing"
output:
  html_document:
    code_folding: show
    df_print: paged
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: yes
editor_options:
  chunk_output_type: inline
---


```{r install, eval=FALSE, include=FALSE}
# parallelization
install.packages("future")
install.packages("furrr")

# timing functions & progress
install.packages("tictoc")
install.packages("microbenchmark")
install.packages("progress")

# advanced example
install.packages("Julia")
```


```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

# Resources

 - https://www.jottr.org/2017/06/05/many-faced-future/
 - https://github.com/HenrikBengtsson/future
 - https://davisvaughan.github.io/furrr/

# Timing functions

```{r}
library(microbenchmark)
tm <- microbenchmark(
  rchisq(100, 0),
  rchisq(100, 1),
  rchisq(100, 2),
  rchisq(100, 3),
  rchisq(100, 5), times = 1000L)
print(tm)
autoplot(tm)
```

```{r}
library(tictoc)
tic()
x <- runif(1e7)
toc()

tic("outer")
x1 <- runif(1e7)
tic("middle")
x2 <- runif(1e7)
tic("inner")
x3 <- runif(1e7)
toc()
toc()
toc()
```

# Setting up multiprocess

```{r}
# general purpose parallelization package
library(future)
availableCores()
plan(multiprocess)
```

# Running map functions in parallel

Note: `purrr` is planning built-in parallelization support in the near future but in the mean time `furrr` provides `future` versions of the most common functions.

```{r}
library(furrr)
df <- 
  data_frame(
    mean = 1:4
  )

# sequential
tic()
df2 <- df %>% 
  mutate(
    norm_sd = map_dbl(mean, ~rnorm(1e7, mean = .x) %>% sd())
  )
toc()

# parallel
tic()
df2 <- df %>% 
  mutate(
    norm_sd = future_map_dbl(mean, ~rnorm(1e7, mean = .x) %>% sd())
  )
toc()
```

## Caveat

Sending large data sets between cores makes even parallel things slower!

```{r}
# sequential
tic()
df2 <- df %>% 
  mutate(
    data = map(mean, ~rnorm(1e7, mean = .x)),
    norm_sd = map_dbl(data, ~sd(.x))
  )
toc()

# parallel
tic()
df2 <- df %>% 
  mutate(
    data = future_map(mean, ~rnorm(1e7, mean = .x)),
    norm_sd = future_map_dbl(data, ~sd(.x))
  )
toc()
```

## Progress Bar

```{r}
# not built into purrr yet but will come
# this type of progress bar is only visible in interactive use (not knitted)
library(progress)
df <- data_frame(mean = 1:100)
pb <- progress_bar$new(
  total = nrow(df), 
  format = "(:spin) [:bar] :percent :elapsed")

tic()
df <- df %>% 
  mutate(
    data = map(mean, ~{
      pb$tick(); rnorm(1e6, mean = .x); .x
    })
  )
toc()
```

```{r}
# parallel processing with furrr (build-in progress bar)
tic()
df %>% 
  mutate(
    data = future_map(mean, ~rnorm(1e6, mean = .x), .progress = TRUE)
  )
toc()
```


# Parallel Fun with Fractals

```{r}
library("Julia")
#' convert julia set output to a tidy data frame
convert_julia_to_data_frame <- function(set) {
  # convert to dplyr data frame
  set %>% as.data.frame.table() %>% tbl_df() %>%
    # rename dimensions and value
    rename(x = Var1, y = Var2, value = Freq) %>%
    # convert dimensions to numbers
    mutate_if(is.factor, as.integer)
}
#' plot julia set
plot_julia_data_frame <-
  function(julia_df, colors = topo.colors(16)) {
    julia_df %>%
      ggplot() +
      aes(x, y, fill = value) +
      geom_raster() +
      scale_fill_gradientn(colors = colors) +
      coord_equal() +
      scale_x_continuous(expand = c(0, 0)) +
      scale_y_continuous(expand = c(0, 0))
  }
```

## Example

```{r}
library(tictoc)
tic()
set <- JuliaImage(500, centre = 0 + 0i, L = 3.5, C = -0.4 + 0.6i)
set %>% convert_julia_to_data_frame() %>% plot_julia_data_frame()
toc()
```

## Different Fractals

```{r}
Cs <- c(
  a = -0.618,
  b = -0.4     + 0.6i,
  c =  0.285   + 0i,
  d =  0.285   + 0.01i,
  e = -0.70176 - 0.3842i,
  f =  0.835   - 0.2321i,
  g = -0.8     + 0.156i,
  h = -0.7269  + 0.1889i,
  i =          - 0.8i
)
Cs
```

```{r}
# sequential
tic()
pset <- 
  Cs %>%
  map(~JuliaImage(500, centre = 0 + 0i, L = 3.5, C = .x))
toc()
```

```{r}
library(future)
library(furrr)
# parallel
tic()
pset <- 
  Cs %>%
  future_map(~JuliaImage(500, centre = 0 + 0i, L = 3.5, C = .x), .progress = TRUE)
toc()
```

### Visualize

```{r}
plot_df <-
  data_frame(
    set = names(pset),
    data = map(pset, convert_julia_to_data_frame)
  ) 
plot_df
```

```{r, fig.width=12, fig.height=12}
plot_df %>% unnest(data) %>% plot_julia_data_frame() + facet_wrap(~set)
```

